---
output:
    github_document:
        df_print: kable
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- knit with rmarkdown::render("README.Rmd", output_format = "md_document") -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- <\!-- badges: start -\-> -->
<!--   [![R-CMD-check](https://github.com/ph-rast/dcnet/workflows/R-CMD-check/badge.svg)](https://github.com/ph-rast/dcnet/actions) -->
<!-- [![R-CMD-check](https://github.com/ph-rast/dcnet/workflows/R-CMD-check/badge.svg)](https://github.com/ph-rast/dcnet/actions) -->
<!-- <\!-- badges: end -\-> -->

## Installation
This package is in development. The current version can be insalled with
```{r}
if (!requireNamespace("remotes")) { 
  install.packages("remotes")   
}   
remotes::install_github("ph-rast/dcnet")
```

## Example: Dynamic Conditional Networks with EMA Data
This section illustrates the use with a dataset that contains synthetic data from an ecological momentary assessment. The data is derived from a 100 Day study conducted at UC Davis (O'Laughlin et al. 2020, Willams et al. 2020).


After loading the package, the `ema_fitbit` data is available as a list of 38 matrices. Each matrix contains the four variables of interest: `active`, `excited`, `totalDistance`, and `interested`. The rows represent 98 daily observations, and the columns correspond to these four variables. Each matrix is for one person (N=38).
```{r}
library(dcnet)

## Head of the data matrix for the first person:
head(ema_fitbit[[1]])
```

Illustrate first person's data. Given that `ema_fitbit` is a list, we first need to extract person 1 and add a time index for `ggplot`:
```{r}
library(ggplot2)

ema_fitbit
df1 <- data.frame(ema_fitbit[[1]], time = 1:98)

## Manually reshape the data to long format using base R
df1_long <- do.call(rbind, lapply(names(df1)[-5], \(var) {
  data.frame(time = df1$time, variable = var, value = df1[[var]])
}))

## Convert the 'variable' column to a factor for consistent ordering in facets
df1_long <- df1_long |>
  within({ variable <- factor(variable, levels = names(df1)[-5]) })

## Plot with ggplot2
plt <- ggplot(df1_long, aes(x = time, y = value)) +
  geom_smooth(show.legend = FALSE, se = FALSE) +
  geom_point(alpha = 0.4) +
  facet_wrap(~ variable, scales = "free_y") +
  labs(x = "Time", y = "Value", title = "Time Series for Variables for Person 1") +
  theme_minimal()

plt
```

## Fit Dynamic Conditional Correlation (DCC) Network Model
Next we will fit the multilevel DCC model for the shocks, the residual conditional covariances, and a multilevel VAR structure for means structure.

The function `dcnet` reads the data in the structure given in `ema_fitbit`. The `parameterization = "DCCr"` selects the DCC with random effects on all components, the garch parameters for the standard deviations as well as all garch parameters for the conditional correlations.
In order to shorten computation time, we will use `sampling_algorithm = 'variational'`. `variational` employs Stan's variational inference algorithm that approximates the posterior via the "meanfield" algorithm. For full sampling, one could use `hmc`, but this option will need to run hours and days to converge. The `init = 0.5` argument is optional, but works well to create good starting values that don't stray too far.  

## Fitting the Multilevel DCC Model

Next, we will fit the **multilevel DCC model** to estimate the shocks, the residual conditional covariances, and a **multilevel VAR structure** for the mean.

The `dcnet()` function reads the data provided in the structure of `ema_fitbit`. Setting the argument `parameterization = "DCCr"` selects the **DCC model with random effects** on all components. This includes random effects for both the **GARCH parameters** governing the standard deviations and the **GARCH parameters** for the conditional correlations.

To shorten computation time, we will use `sampling_algorithm = "variational"`. The `variational` algorithm employs **Stan's variational inference** to approximate the posterior distribution using the **meanfield** approximation. For full Bayesian sampling, one could set `sampling_algorithm = "hmc"`. However, this option is computationally expensive and may take several hours or even days to converge.

The argument `init = 0.5` is optional but often useful, as it provides reasonable starting values that help the algorithm converge faster without straying too far from the posterior space.
```{r}
fit <- dcnet(data = ema_fitbit, parameterization = "DCCr",
             sampling_algorithm = 'variational', init = 0.5)
```

Once the model is fit, we can request a summary (check your RAM -- this step might kill R if you are close to maxing out your memory):
```{r}
summary(fit)
```


## References
- O'Laughlin, K. D., Liu, S., & Ferrer, E. (2020). Use of Composites in Analysis of Individual
Time Series: Implications for Person-Specific Dynamic Parameters. Multivariate
Behavioral Research, 1–18. https://doi.org/10.1080/00273171.2020.1716673
- Williams, D. R., Martin, S. R., Liu, S., & Rast, P. (2020). Bayesian Multivariate
Mixed-Effects Location Scale Modeling of Longitudinal Relations Among Affective
Traits, States, and Physical Activity. European Journal of Psychological
Assessment, 36 (6), 981–997. https://doi.org/10.1027/1015-5759/a000624
